\section{Introduction}
\subsection{Motivation}
This project was initially motivated by my lived experience as a \gls{neurodivergent} developer and by personal challenges in using industrial software development tools. In professional settings, my \gls{ASD} was 
often framed as a ``superpower'' due to my ability to rapidly acquire new technical skills and engage in deep logical analysis. However, this perceived strength came with a significant \gls{cognitive_tax}. I found adapting
 to new environments and maintaining a sustainable work-life balance extremely difficult, particularly when faced with task initiation and the decomposition of complex programming problems.
Upon researching existing methodologies and consulting with other neurodivergent developers, it became clear that current Integrated Development Environments (IDEs) offer no effective, ``out-of-the-box'' solutions
 for managing these executive functioning barriers. This gap in assistive technology motivated the development of a solution that transitions the IDE from a passive tool to an active cognitive partner.

Existing \glspl{IDE} can lead to increased cognitive overload and difficulty in maintaining focus. By streamlining interfaces and providing \gls{AI}-driven task decomposition, this project aims 
to reduce the strain on \gls{neurodivergent} individuals in both professional and educational settings, as developers often report challenges with cognitive overload, \gls{hyperfocus}, and breaking
 down more complex problems when programming

\subsection{Problem Statement}
\glspl{IDE} are a primary coding tool for modern software development, yet there are still elements of the interface that remain a significant barrier for \gls{neurodivergent} developers, specifically
 those with \gls{ADHD}, \gls{ASD}, and \gls{dyslexia}. While these tools are feature-rich, they are designed with a ``neurotypical bias'' that assumes a high baseline for filtering distractions and
  managing multi-step planning.

Current IDEs prioritize high feature density, often presenting a constant stream of visual stimuli, including nested menus, status bars, and dense technical jargon. For developers who experience 
sensory overwhelm, this environment creates an immense Extraneous Cognitive Load~\cite{sweller1988cognitive}. This ``visual noise'' competes for limited working memory, often exhausting a developer’s
 ``cognitive budget'' before they can engage with the actual logic of the code.

Research indicates that developers with \gls{ADHD} frequently struggle with task paralysis during the initiation phase of a project~\cite{bramham2009executive}. Furthermore, those with \gls{ASD} may face 
challenges with high-level strategic planning and the decomposition of abstract requirements into actionable steps. Constant notifications—such as linter warnings or compiler recommendations—can inadvertently trigger hyperfocus on peripheral
  details, leading to ``rabbit holes'' that derail the core development goal.

While AI tools are increasingly used in software development, they often focus on generating code rather than supporting the process of being a developer. There is a lack of ``out-of-the-box'' inclusivity in 
these tools, forcing neurodivergent individuals to manually create complex custom configurations to remain productive. Consequently, there is a critical need for an IDE intervention that acts as a 
``digital ramp'', one that filters environmental noise and provides the necessary scaffolding to bypass the freeze response common in task paralysis.

\subsection{Understanding Neurodiversity in Software Development}
Before going into the specific technical barriers of modern \glspl{IDE}, it is necessary to define neurodiversity and the cognitive
 frameworks that underpin this research. Originally coined by Judy Singer in the late 1990s, ``neurodiversity''  views neurological differences—such 
 as \gls{ASD}, \gls{ADHD}, and \gls{dyslexia}, as natural variations in the human genome rather than purely medical deficits~\cite{singer1999why}.

In the context of software engineering, these variations often manifest as unique cognitive processing styles. While many \gls{neurodivergent} 
developers possess high aptitude for logic and pattern recognition, they frequently encounter barriers related to \gls{executive_dysfunction}. 
This is an umbrella term for difficulty in mental processes that enable us to plan, focus attention, remember instructions, and juggle multiple tasks 
successfully. For a developer, executive dysfunction is not a lack of technical ability, but rather a barrier in three key areas: Inhibitory Control
 (filtering IDE ``noise''), Working Memory (tracking complex syntax), and Cognitive Flexibility (switching between high-level logic and low-level code).

This challenge is further complicated by the ``Double Empathy'' problem—the idea that communication breakdowns occur because different groups process information differently, rather than one group being
 inherently ``wrong''. In the context of this project, neurodivergent developers may struggle with ``implied'' logic in a team setting, where their interpretation of a task differs from the team's original
  plan. However, if instructions are provided clearly and literally, these challenges can be mitigated.
 Furthermore, while neurodivergent developers often utilize hyperfocus to solve complex problems, this state is inherently fragile. A single interruption—such as a non-essential notification—can lead
  to a total loss of mental context, carrying a ``state recovery cost'' that significantly hinders productivity. By identifying these specific cognitive roadblocks, we can move away from a
    medical model of disability and toward a social model that adapts the development environment to the user's brain.


To provide a clearer lens for the objectives of this project, the below table outlines the specific strengths and environmental challenges associated with the conditions most prevalent in the developer community.
\begin{table}[H]
    \centering
    \caption{Neurodivergent Strengths and Environmental Challenges}
    \renewcommand{\arraystretch}{1.5} % Adds breathing room between rows
    \begin{tabularx}{\textwidth}{l >{\raggedright\arraybackslash}X >{\raggedright\arraybackslash}X}
        \toprule
        \textbf{Condition} & \textbf{Potential Strengths} & \textbf{Common Environmental Challenges} \\
        \midrule
        \gls{ASD} & Deep focus, systematic thinking, high integrity, exceptional memory for details, strong pattern recognition. & Sensory overload in noisy or bright environments, misinterpretation of direct communication style, difficulty with unspoken social rules. \\
        \addlinespace
        \gls{ADHD} & Creativity, hyperfocus on passions, adaptability, high energy, ability to multitask or rapidly switch contexts. & Distractibility in chaotic settings, struggles with time management and organisation, impulsivity in structured environments. \\
        \addlinespace
        Dyslexia & Strong visual-spatial reasoning, excellent big-picture thinking, creative problem-solving, strong verbal skills. & Difficulties with text-heavy tasks, timed reading assignments, and conventional written communication. \\
        \bottomrule
    \end{tabularx}
\end{table}

\subsection{Aims and Objectives}

\textbf{Project Aim}

The primary objective of this project is to design, implement, and evaluate an assistive VS Code plug-in that reduces extraneous cognitive load and mitigates the effects of
 \gls{executive_dysfunction} for \gls{neurodivergent} software developers. By introducing a ``cognitive prosthetic'' layer, the project seeks
  to determine if targeted \gls{UI} simplification and \gls{AI}-driven task scaffolding can measurably improve the developer experience.

\textbf{Research Objectives}

\textbf{Objective 1:} Interface Optimisation (The Minimalist Toggle)

Target: Develop a single ``Minimalist Toggle'' feature that simultaneously hides non-essential \gls{UI} elements, including the Activity Bar, Status Bar, Mini-map, and Breadcrumbs.


Justification: This directly addresses the issue of visual overload by reducing the number of stimuli competing for the developer's attention, thus lowering the Extraneous
 Cognitive Load and creating a more focused coding environment \cite{sweller1988cognitive}, reducing visual noise that leads to distraction and sensory overwhelm for autistic and \gls{ADHD} developers.

\textbf{Objective 2: }\gls{AI}-Driven Task Decomposition (The Task Architect)

Target: To implement an \gls{LLM}-integrated module that translates ambiguous programming goals into structured, editable checklists using a Chain of Thought (CoT) prompting strategy. This should be
implemented via the Open\gls{AI} \gls{API} that breaks down high-level requirements without providing the direct code solution \cite{ringel2019ai}.


Justification: This supports ``Task Initiation'' and reduces the energy required to overcome task paralysis, a primary barrier for developers with ADHD and ASD \cite{gama2025socio}.

\textbf{Objective 3:} Simplification of Technical Feedback (The Code Explainer)

Target: To build a diagnostic interceptor that re-interprets complex compiler errors into plain-language, bulleted instructions~\cite{rocha2025conceptual}. Limit output to a maximum of three bullet points using a ``Direct Instruction'' style.


Justification: This minimizes the cognitive demand of interpreting dense jargon, preventing ``rabbit holes'' and supporting the literal communication needs identified in the Double Empathy
 problem. By simplifying the feedback loop, developers can maintain focus on the core problem rather than getting lost in peripheral details. This is particularly beneficial for those with ASD,
  who may struggle with implied logic and prefer clear, direct instructions. Additionally, it can help mitigate the hyperfocus trap by providing concise guidance that keeps the developer on track 
  without overwhelming them with information.

\textbf{Objective 4: }Visual Accessibility Customisation (The Visual Styler)

Target: To create at least two visual presets, such as ``Soft High Contrast'', that adhere to WCAG 2.1 guidelines to reduce eye strain and enhance readability~\cite{bengtsson2025web}.

Justification: Addresses the sensory sensitivities and ``attractive colour'' preferences reported by \gls{neurodivergent} developers in research surveys~\cite{rocha2025conceptual}.

\textbf{Objective 5:} Empirical Evaluation of Cognitive Impact

Target: To conduct a comparative study using the \gls{NASATLX} and the \gls{SUS} to validate the plugin's efficacy against a standard IDE.

Justification: This provides a standardized scientific metric to confirm if the tool successfully reduces mental demand and frustration for the target demographic.


