\textbf{Abstract}
Modern Integrated Development Environments (\glspl{IDE}) prioritize feature density, often at the expense of cognitive accessibility. 
For \gls{neurodivergent} software developers—particularly those with \gls{ADHD}, \gls{ASD}, or dyslexia—this complexity manifests as 
significant barriers driven by \gls{executive_dysfunction}, including cognitive overload and difficulty in task initiation. 
This research investigates the efficacy of an \gls{AI}-integrated plugin designed to mitigate these challenges within the VS Code 
ecosystem.


The developed solution comprises three core modules: a toggle to close all unnessary panels to filter interface-driven distractions, 
a tack breakdown tool to leverage Large Language Models (\glspl{LLM}) to decompose complex requirements into structured checklists, 
and a visual styling tool for high-contrast, low-strain environmental presets. The system's impact was evaluated through a comparative
 user study ($N=5$) pitting a ``Vanilla'' VS Code baseline against the enhanced environment. Evaluation metrics included the 
 \gls{NASATLX} \cite{NASATLX} to quantify shifts in mental workload and the \gls{SUS} \cite{Hyzy2022} to assess overall usability 
 and tool effectiveness.
 
 Results indicate that by streamlining the development environment and providing automated task scaffolding, 
 the plugin significantly reduces the ``executive tax'' imposed by standard tools. This allows \gls{neurodivergent} developers to 
 reallocate critical cognitive resources from interface navigation to core algorithmic problem-solving.